---
title: "ELF-MERGE"
output: html_notebook
editor_options: 
  chunk_output_type: console
---


Last updated 29/06/23 06:28

# libraries and folders


```{r}




library(sf)
library(tidyverse)
# library(mapview)
library(cowplot)
library(lubridate)
# library(xlsx)
library(readxl)
library(rmapshaper)
library(mapview)
# library(rmapshaper)
library(plotly)
library(tictoc)
library(parallel)
library(doParallel)
library(terra)
library(openxlsx)
library(scales)
library(purrr)
library(furrr)
library(terra)

Current_PC <- Sys.info()["nodename"]

if (Current_PC == ""){
  root = "T:/Palmerston North/Projects A-E/Environmental Limiting Factors/ELF_Project/"
  datadir = paste0(root, "DATA/")
  plotdir = paste0(root, "plot/")


} else if (Current_PC == "EANZ-DT01-Linux") {
  os <- "linux"
  rootDrive <- "/media/andrew/HDD data/"
  EDrive <- "/media/andrew/E-DRIVE-SSD/"
  GeoSpatRoot = "C:/Users/McMillanAn/OneDrive - MWLR/NZ-Geospatial-Data/"
  
} else if (Current_PC == "EANZ-DT01") {
  os <- "win10"
  rootDrive <- "D:/"
  EDrive <- "E:/"
  GeoSpatRoot = "C:/Users/McMillanAn/OneDrive - MWLR/NZ-Geospatial-Data/" #check this!
  
} else if (Current_PC == "styx") {
  os <- "linux"
  rootDrive <- "/home/mcmillana/"
  elfdir = "/mnt/data/mcmillana/PRJ3597-MPI-ELF/"
  GeoSpatRoot = "/mnt/data/mcmillana/NZ-Geospatial-Data/"
  source("/home/mcmillana/code/R/utils/mcm_sp_utils.R") 
  
} else if (Current_PC == "L-8P30JW3") {
  os <- "win10"
  rootDrive <- "C:/Users/McMillanAn/OneDrive - MWLR"
  elfdir = "T:/Palmerston North/Projects A-E/Environmental Limiting Factors/ELF_Project/"
  GeoSpatRoot = "C:/Users/McMillanAn/OneDrive - MWLR/NZ-Geospatial-Data/"
  source("T:/Palmerston North/Projects A-E/Environmental Limiting Factors/ELF_Project/DATA/SOIL_Limitation_layers/Rcode/mcm_sp_utils.R")  
} else {
  rootDrive <- "D:/"
}


# NOTE CORRECTING LAYER 3/7/23
elf_merge_dir = paste0(elfdir, "DATA/MERGING_ELF_REV/")
if(!dir.exists(elf_merge_dir)){dir.create(elf_merge_dir)}
```

# function: find_overlapping 

```{r}

merge_elf_layers = function(LA_INPUT, LB_INPUT, LB_LIM_FIELDNAME, LB_LIMITATION_descr, elf_merge_dir,  ffn_stub_list, ncore_rec = 36){
  
  #----test input data
  # elf_merge_dir = paste0(elfdir, "DATA/MERGING_ELF/")
  
  
  
  testmode = F
  
  if (testmode){
  #Merge PODZOLS into ELF_TO_KEEP to create the merged layer "L0_L1_MERGED"
 
#    LIMITING_SOILS_PODZOLS_ffn = paste0(elfdir,"DATA/SOIL_Limitation_layers/finished_unmerged_layers/LIMITING_SOILS_PODZOLS.shp")
# LIMITING_SOILS_PODZOLS = st_read(LIMITING_SOILS_PODZOLS_ffn)
# 
# ELF_2012_TO_KEEP_PROPER_ffn = paste0(elf_merge_dir,"ELF_2012_TO_KEEP_PROPER.shp")
# ELF_2012_TO_KEEP_PROPER = st_read(ELF_2012_TO_KEEP_PROPER_ffn)
# 
# 
#   ffn_stub_list = c("L0xL1", "L0_minus_L0xL1", "L1_minus_L0xL1", "L0_L1_MERGED")
#   LA_INPUT = ELF_2012_TO_KEEP_PROPER
#   LB_INPUT = LIMITING_SOILS_PODZOLS
#   LB_LIM_FIELDNAME = "PODZOLS"
#   LB_LIMITATION_descr = "Substrate Podzols"
#   
  #Merge TREELINE into L0_L1_MERGED to create the merged layer "L1_L2_MERGED"

  ffn_stub_list = c("L1xL2", "L1_minus_L1xL2", "L2_minus_L1xL2", "L1_L2_MERGED")
  LA_INPUT = st_read(paste0(elf_merge_dir, "L0_L1_MERGED.gpkg"))
  LB_INPUT = st_read(paste0(elf_merge_dir, "NZ-High-Alt-Treeline-layer-02_23_v2.gpkg"))
  mapview(LB_INPUT)
  LB_LIM_FIELDNAME = "TREELINE"
  LB_LIMITATION_descr = "Too Cold Near Treeline"

   #Merge TOO_WET into L1_L2_MERGED to create the merged layer "L1_L2_MERGED"
  
  # ffn_stub_list = c("L2xL3", "L3_minus_L2xL3", "L2_minus_L2xL3", "L2_L3_MERGED")
  # LA_INPUT = st_read(paste0(elf_merge_dir, "L1_L2_MERGED.gpkg"))
  # LB_INPUT = st_read(paste0(elfdir,"DATA/SOIL_Limitation_layers/finished_unmerged_layers/SOIL_LIMITATIONS_TOO_WET.shp"))
  # mapview(LB_INPUT)
  # LB_LIM_FIELDNAME = "WET"
  # LB_LIMITATION_descr = "Too Wet"
  
  ncore_rec = 36
  
  }
  
  
 
  #------------------
  
  tic("whole function")
  LA_ORIG =  LA_INPUT %>%  mutate(LA_UID = 1:nrow(.))
  LB_ORIG =  LB_INPUT %>%  mutate(LB_UID = 1:nrow(.))
  
  
  
  LA = LA_ORIG  %>% select(LA_UID, geometry=contains("geom"))
  LB = LB_ORIG  %>% select(LB_UID, geometry=contains("geom"))
  
  # if (LB_LIM_FIELDNAME == "TREELINE"){
  #     LB = LB_ORIG %>% select(LB_UID, geometry = geom)
  #   }else{
  #     LB = LB_ORIG %>% select(LB_UID, geometry)
  #   }
  
  LA_x_LB = st_intersects(LA,LB)
  length(LA_x_LB)
  
  LA_overlaps_ix = which(lengths(LA_x_LB) > 0)
  length(LA_overlaps_ix)
  LA_overlaps = LA[LA_overlaps_ix,]
  
  # tic()
  # P2_x_P1 = st_intersects(P2,P1_overlaps)
  # toc()
  # length(P2_x_P1)
  
  print("~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~")
  print("++++++++")
  print("++++++++   merge_elf_layers     ++++++ ---> Spatial operation 1 of 3:  Intersection: LAxLB")
  print("++++++++")
  print("~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~")
  
  LAxLB = ea_intersection(LB, LA_overlaps,PC = Current_PC, ffn_stub = ffn_stub_list[1], dn = elf_merge_dir, parallel = T, ncore = ncore_rec, val = "both",diff_opt = F, filetype = "SHP" )
  
  LAxLB_unn = st_union(LAxLB)
  
  print("~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~")
  print("++++++++")
  print("++++++++   merge_elf_layers     ++++++ ---> Spatial operation 2 of 3:  Differencing: LA_minus_LAxLB")
  print("++++++++")
  print("~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~")
  
  LA_minus_LAxLB = ea_intersection(LA, LAxLB_unn,PC = Current_PC, ffn_stub = ffn_stub_list[2], dn = elf_merge_dir, parallel = T, ncore = ncore_rec, val = "P1",filetype = "SHP",diff_opt = T )
  
  print("~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~")
  print("++++++++")
  print("++++++++   merge_elf_layers     ++++++ ---> Spatial operation 3 of 3:  Differencing: LB_minus_LAxLB")
  print("++++++++")
  print("~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~")
  
  
  LB_minus_LAxLB = ea_intersection(LB, LAxLB_unn,PC = Current_PC, ffn_stub = ffn_stub_list[3], dn = elf_merge_dir, parallel = T, ncore = ncore_rec, val = "P1",filetype = "SHP",diff_opt = T )
  
  # Step 4a - Limitation attributes for LAxLB - inherit all from LA, add new from LB and change type to "Multiple"
  LA_attr = LA_ORIG %>% st_set_geometry(NULL) %>% as_tibble()
  
  
  LAxLB_w_attr = LAxLB %>% 
    left_join(LA_attr, by = "LA_UID") %>% 
    mutate(
      !!LB_LIM_FIELDNAME := 1,
      LIMITATION = "Multiple")
  
  # Step 4b - Limitation attributes for LA_minus_LAxLB - inherit all attr and type from LA
  
  LA_minus_LAxLB_w_attr = LA_minus_LAxLB %>% 
    left_join(LA_attr, by = "LA_UID") 
  
  # Step 4c - Limitation attributes for LB_minus_LAxLB - this will have a single limitation which will be the input layer LB
  
  LB_minus_LAxLB_w_attr = LB_minus_LAxLB %>% 
    mutate(
      DRY=0, 
      LIMITED=0,
      FROST_FLAT=0,
      SALINE=0,
      ULTRAMAFIC=0,
      HOT_GEOTHM=0,
      PAKIHI=0,
      GUMLANDS=0,
      SALT_SPRAY=0,
      LIMITATION=0,
      WET=0,
      COLD_BASIN=0, 
      PODZOLS =0,
      TREELINE=0) %>% 
    mutate(
      LIMITED = 1,
      !!LB_LIM_FIELDNAME := 1,
      LIMITATION = LB_LIMITATION_descr
    ) 
    
  # mapview(LAxLB_w_attr, zcol = "LIMITATION", alpha =0 )
  # mapview(LA_minus_LAxLB_w_attr, zcol = "LIMITATION", alpha =0 )
  # mapview(LB_minus_LAxLB_w_attr, zcol = "LIMITATION", alpha =0 )
  
  # Step 5 - Bind the three sets of data together, remove the UID and save.
  
  LA_LB_MERGED = bind_rows(LAxLB_w_attr, LA_minus_LAxLB_w_attr, LB_minus_LAxLB_w_attr) %>% 
    select(!contains("UID"))

  # names(LA_LB_MERGED)
  
  # LA_LB_MERGED %>% filter(LIMITATION == "Multiple" & PODZOLS == 1)
  
  Length_non_valid = length(which(!st_is_valid(LA_LB_MERGED)))
  
  AREA_LA = calc_tot_area(LA)/1e3
  AREA_LB = calc_tot_area(LB)/1e3
  AREA_LAxLB = calc_tot_area(LAxLB)/1e3
  AREA_LA_minus_LAxLB = calc_tot_area(LA_minus_LAxLB)/1e3
  AREA_LB_minus_LAxLB = calc_tot_area(LB_minus_LAxLB)/1e3
  
  AREA_MRG_TH = AREA_LAxLB + AREA_LA_minus_LAxLB + AREA_LA_minus_LAxLB
  AREA_MRG_AC = calc_tot_area(LA_LB_MERGED)/1e3
  
  LA_LB_MERGED_ffn = paste0(elf_merge_dir, ffn_stub_list[[4]], ".gpkg")
  st_write(LA_LB_MERGED, LA_LB_MERGED_ffn, append = F)
  print(paste("Writing the Newly Merged layer to", LA_LB_MERGED_ffn))
  mapview(LA_LB_MERGED)
  # Step 5b - Replace 
  
  # Check nrows
  # LA_LB_MERGED = st_read(LA_LB_MERGED_ffn)  
  # LA_LB_MERGED %>% filter(LIMITATION == "Multiple" & PODZOLS ==1 ) %>%  nrow()
  # 
  
  
  WF = toc()
  print(paste("whole merge function took",WF))
  
  OUT = list(LAxLB, LA_minus_LAxLB, LB_minus_LAxLB, LA_LB_MERGED )
}

```

LAYERS

L0 : ELF_2012_KEEP        : /DATA/SOIL_Limitation_layers/finished_unmerged_layers/ELF_2012_TO_KEEP_PROPER.shp
L1 : PODZOLS              : /DATA/SOIL_Limitation_layers/finished_unmerged_layers/LIMITING_SOILS_PODZOLS.shp
L2 : TREELINE             : /DATA/SOIL_Limitation_layers/finished_unmerged_layers/NZ-High-Alt-Treeline-layer-02_23_v2.gpkg
L3 : TOO_WET              : /DATA/SOIL_Limitation_layers/finished_unmerged_layers/LIMITING_SOILS_TOO_WET.shp
L4 : TOO_DRY              : /DATA/SOIL_Limitation_layers/finished_unmerged_layers/LIMITING_SOILS_TOO_DRY.shp
L5 : SALINE               : /DATA/SOIL_Limitation_layers/finished_unmerged_layers/SMAP_FSL_CMB_NZSC_FILT_SALINE.shp

Approach for merging a new layer Lb into the exisiting layer La

1. LaxLb: Get Intersection LaxLb. These are the polygons that will have a minimum of two limiting factors (at least one from La and exactly one from Lb)
2. La_minus_LaxLb: Get the difference between La and LaxLb. These are the polygons from the old layer that remain
3. Lb_minus_LaxLb: Get the difference between Lb and LaxLb. These are the polygons from the new layer that will be merged and extend the area of limitation

# merging layers

```{r}


elf_2012_layer_ffn = "/mnt/data/mcmillana/PRJ3597-MPI-ELF/DATA/mfe-environmental-limiting-factors-2012-SHP/environmental-limiting-factors-2012.shp"
elf_2012_layer=st_read(elf_2012_layer_ffn)

names(elf_2012_layer)
elf_2012_layer %>% pull(LIMITATION) %>% unique()

# [1] "DRY"        "LIMITED"    "FROST_FLAT" "SALINE"     "ULTRAMAFIC" "HOT_GEOTHM" "PAKIHI"     "GUMLANDS"   "SALT_SPRAY" "LIMITATION"
# [11] "WET"        "COLD_BASIN" "PODZOLS"    "TREELINE"   "geometry"  

# > elf_2012_layer %>% pull(LIMITATION) %>% unique()

#  [1] "None"                   "Substrate Gumlands"     "Substrate Pakihi"       "Substrate Podzols"      "Too Exposed Salt Spray"
#  [6] "Substrate Ultramafic"   "Multiple"               "Too Cold Frost Flats"   "Too Wet"                "Too Cold Near Treeline"
# [11] "Too Dry"                "Too Hot Geothermal"     "Too Cold Basins"   





LIMITING_SOILS_PODZOLS_ffn = paste0(elfdir,"DATA/SOIL_Limitation_layers/finished_unmerged_layers/LIMITING_SOILS_PODZOLS.shp")
LIMITING_SOILS_PODZOLS = st_read(LIMITING_SOILS_PODZOLS_ffn)

ELF_2012_TO_KEEP_PROPER_ffn = paste0(elf_merge_dir,"ELF_2012_TO_KEEP_PROPER.shp")
ELF_2012_TO_KEEP_PROPER = st_read(ELF_2012_TO_KEEP_PROPER_ffn)



L0 = ELF_2012_TO_KEEP_PROPER
L1 = LIMITING_SOILS_PODZOLS

LB_LIM_TYPE = "PODZOLS"
LB_LIMITATION_descr = "Substrate Podzols"
ffn_stub_curr = "L0xL1"

ffn_stub_list = c("L0xL1", "L0_minus_L0xL1", "L1_minus_L0xL1")

MRG_LAYERS = merge_elf_layers(L0,L1, datadir = elf_merge_dir, ffn_stub_list = ffn_stub_list,ncore_rec = 36) #551 seconds

MRG_LAYERS = L0xL1

MRG_LAYERS[[1]]
MRG_LAYERS[[2]]
MRG_LAYERS[[3]]


L0_L1_MERGED = st_read(paste0(elf_merge_dir, "L0_L1_MERGED.gpkg"))

L0_L1_MERGED_f = L0_L1_MERGED %>% filter(LIMITATION=="Multiple" & PODZOLS==1)

```




# MERGE_L0_L1 : PODZOLS (L1) ---> ELF_TO_KEEP_PROPER (L0)

```{r}

   LIMITING_SOILS_PODZOLS_ffn = paste0(elfdir,"DATA/SOIL_Limitation_layers/finished_unmerged_layers/LIMITING_SOILS_PODZOLS.shp")
LIMITING_SOILS_PODZOLS = st_read(LIMITING_SOILS_PODZOLS_ffn)

ELF_2012_TO_KEEP_PROPER_ffn = paste0(elf_merge_dir,"ELF_2012_TO_KEEP_PROPER.shp")
ELF_2012_TO_KEEP_PROPER = st_read(ELF_2012_TO_KEEP_PROPER_ffn)


  ffn_stub_list = c("L0xL1", "L0_minus_L0xL1", "L1_minus_L0xL1", "L0_L1_MERGED")
  LA_INPUT = ELF_2012_TO_KEEP_PROPER
  LB_INPUT = LIMITING_SOILS_PODZOLS
  LB_LIM_FIELDNAME = "PODZOLS"
  LB_LIMITATION_descr = "Substrate Podzols"

  MERGE_L0_L1_RES = merge_elf_layers(LA_INPUT, LB_INPUT, LB_LIM_FIELDNAME, LB_LIMITATION_descr, elf_merge_dir,  ffn_stub_list, ncore_rec = 36) 
  
```
 

# MERGE_L1_L2 : TREELINE (L2) ---> L0_L1_MERGED 

```{r}

   #Merge TREELINE into L0_L1_MERGED to create the merged layer "L1_L2_MERGED"

  ffn_stub_list = c("L1xL2", "L1_minus_L1xL2", "L2_minus_L1xL2", "L1_L2_MERGED")
  LA_INPUT = st_read(paste0(elf_merge_dir, "L0_L1_MERGED.gpkg"))
  LB_INPUT = st_read(paste0(elf_merge_dir, "NZ-High-Alt-Treeline-layer-02_23_v2.gpkg"))
  # mapview(LB_INPUT)
  LB_LIM_FIELDNAME = "TREELINE"
  LB_LIMITATION_descr = "Too Cold Near Treeline"

  MERGE_L1_L2_RES = merge_elf_layers(LA_INPUT, LB_INPUT, LB_LIM_FIELDNAME, LB_LIMITATION_descr, elf_merge_dir,  ffn_stub_list, ncore_rec = 64)

```
 
 

# MERGE_L2_L3 : WET (L3) ---> L1_L2_MERGED 

```{r}
#Merge TOO_WET into L1_L2_MERGED to create the merged layer "L2_L3_MERGED"
  
  ffn_stub_list = c("L2xL3", "L2_minus_L2xL3", "L3_minus_L2xL3", "L2_L3_MERGED")
  LA_INPUT = st_read(paste0(elf_merge_dir, "L1_L2_MERGED.gpkg"))
  LB_INPUT = st_read(paste0(elfdir,"DATA/SOIL_Limitation_layers/finished_unmerged_layers/SOIL_LIMITATIONS_TOO_WET.shp"))
  mapview(LB_INPUT)
  LB_LIM_FIELDNAME = "WET"
  LB_LIMITATION_descr = "Too Wet"

  MERGE_L2_L3_RES = merge_elf_layers(LA_INPUT, LB_INPUT, LB_LIM_FIELDNAME, LB_LIMITATION_descr, elf_merge_dir,  ffn_stub_list, ncore_rec = 36)

```
 

# MERGE_L3_L4 : DRY (L4) ---> L2_L3_MERGED 

```{r}
#Merge TOO_DRY into L2_L3_MERGED to create the merged layer "L3_L4_MERGED"
  
  ffn_stub_list = c("L3xL4", "L3_minus_L3xL4", "L4_minus_L3xL4", "L3_L4_MERGED")
  LA_INPUT = st_read(paste0(elf_merge_dir, "L2_L3_MERGED.gpkg"))
  LB_INPUT = st_read(paste0(elfdir,"DATA/SOIL_Limitation_layers/finished_unmerged_layers/SOIL_LIMITATIONS_TOO_DRY.shp"))
  # mapview(LB_INPUT)
  LB_LIM_FIELDNAME = "DRY"
  LB_LIMITATION_descr = "Too Dry"

  MERGE_L3_L4_RES = merge_elf_layers(LA_INPUT, LB_INPUT, LB_LIM_FIELDNAME, LB_LIMITATION_descr, elf_merge_dir,  ffn_stub_list, ncore_rec = 36)
  
  L3_L4_MERGED = st_read(paste0(elf_merge_dir, "L3_L4_MERGED.gpkg"))
  mapview(L3_L4_MERGED, zcol = "LIMITATION")
  

```
 

# MERGE_L4_L5 : SALINE (L5) ---> L3_L4_MERGED 

```{r}
#Merge SALINE into L3_L4_MERGED to create the merged layer "L4_L5_MERGED"
  
  ffn_stub_list = c("L4xL5", "L4_minus_L4xL5", "L5_minus_L4xL5", "L4_L5_MERGED")
  LA_INPUT = st_read(paste0(elf_merge_dir, "L3_L4_MERGED.gpkg"))
  LB_INPUT = st_read(paste0(elfdir,"DATA/SOIL_Limitation_layers/finished_unmerged_layers/SMAP_FSL_CMB_NZSC_FILT_SALINE.shp"))
  # mapview(LB_INPUT)
  LB_LIM_FIELDNAME = "SALINE"
  LB_LIMITATION_descr = "Substrate Saline"

  MERGE_L4_L5_RES = merge_elf_layers(LA_INPUT, LB_INPUT, LB_LIM_FIELDNAME, LB_LIMITATION_descr, elf_merge_dir,  ffn_stub_list, ncore_rec = 36)

```
 
# Area Checks 
 
```{r}

# read the 2012 Layer
ELF_2012_ffn = paste0(elfdir, "DATA/mfe-environmental-limiting-factors-2012-SHP/environmental-limiting-factors-2012.shp")
ELF_2012 = st_read(ELF_2012_ffn) %>% 
  filter(!LIMITATION=="None")

ELF_2012_ng = ELF_2012 %>% ea_id_area(rem_geo = T) 
ELF_2012_ng_SMY = ELF_2012_ng %>% group_by(LIMITATION) %>% summarise(AREA_KHA = sum(POLY_AREA_HA)/1e3)

ELF_2012_ng %>% select(DRY, FROST_FLAT:SALT_SPRAY, WET:TREELINE, POLY_AREA_HA) %>% 
  gather(LIM, BX, -POLY_AREA_HA) %>% 
  filter(BX==1) %>% 
  group_by(LIM) %>% 
  summarise(AREA_KHA = sum(POLY_AREA_HA)/1e3)
  
  

 LIMITING_SOILS_PODZOLS_ffn = paste0(elfdir,"DATA/SOIL_Limitation_layers/finished_unmerged_layers/LIMITING_SOILS_PODZOLS.shp")
LIMITING_SOILS_PODZOLS = st_read(LIMITING_SOILS_PODZOLS_ffn)

ELF_2012_TO_KEEP_PROPER_ffn = paste0(elf_merge_dir,"ELF_2012_TO_KEEP_PROPER.shp")
ELF_2012_TO_KEEP_PROPER = st_read(ELF_2012_TO_KEEP_PROPER_ffn)

calc_tot_area(LIMITING_SOILS_PODZOLS)











```





 
 
  