---
title: "ELF-MERGE"
output: html_notebook
editor_options: 
  chunk_output_type: console
---


Last updated 29/06/23 06:28

# libraries and folders


```{r}




library(sf)
library(tidyverse)
# library(mapview)
library(cowplot)
library(lubridate)
# library(xlsx)
library(readxl)
library(rmapshaper)
library(mapview)
# library(rmapshaper)
library(plotly)
library(tictoc)
library(parallel)
library(doParallel)
library(terra)
library(openxlsx)
library(scales)
library(purrr)
library(furrr)
library(terra)

Current_PC <- Sys.info()["nodename"]

if (Current_PC == ""){
  root = "T:/Palmerston North/Projects A-E/Environmental Limiting Factors/ELF_Project/"
  datadir = paste0(root, "DATA/")
  plotdir = paste0(root, "plot/")


} else if (Current_PC == "EANZ-DT01-Linux") {
  os <- "linux"
  rootDrive <- "/media/andrew/HDD data/"
  EDrive <- "/media/andrew/E-DRIVE-SSD/"
  GeoSpatRoot = "C:/Users/McMillanAn/OneDrive - MWLR/NZ-Geospatial-Data/"
  
} else if (Current_PC == "EANZ-DT01") {
  os <- "win10"
  rootDrive <- "D:/"
  EDrive <- "E:/"
  GeoSpatRoot = "C:/Users/McMillanAn/OneDrive - MWLR/NZ-Geospatial-Data/" #check this!
  
} else if (Current_PC == "styx") {
  os <- "linux"
  rootDrive <- "/home/mcmillana/"
  elfdir = "/mnt/data/mcmillana/PRJ3597-MPI-ELF/"
  elf_merge_dir = paste0(elfdir, "DATA/MERGING_ELF_REV/")
  GeoSpatRoot = "/mnt/data/mcmillana/NZ-Geospatial-Data/"
  source("/home/mcmillana/code/R/utils/mcm_sp_utils.R") 
  
} else if (Current_PC == "L-8P30JW3") {
  os <- "win10"
  rootDrive <- "C:/Users/McMillanAn/OneDrive - MWLR"
  # elfdir = "T:/Palmerston North/Projects A-E/Environmental Limiting Factors/ELF_Project/"
  elfdir = "C:/Users/McMillanAn/OneDrive - MWLR/Projects/PRJ3597-MPI-ELF/"
  elf_merge_dir = paste0(elfdir, "data/MERGING_ELF_REV/")
  GeoSpatRoot = "C:/Users/McMillanAn/OneDrive - MWLR/NZ-Geospatial-Data/"
  source("C:/Users/McMillanAn/OneDrive - MWLR/code/R/mcm_sp_utils.R")  
  
} else {
  rootDrive <- "D:/"
}


# NOTE CORRECTING LAYER 3/7/23

if(!dir.exists(elf_merge_dir)){dir.create(elf_merge_dir)}



```

LAYERS

L0 : ELF_2012_KEEP        : /DATA/SOIL_Limitation_layers/finished_unmerged_layers/ELF_2012_TO_KEEP_PROPER.shp
L1 : PODZOLS              : /DATA/SOIL_Limitation_layers/finished_unmerged_layers/LIMITING_SOILS_PODZOLS.shp
L2 : TREELINE             : /DATA/SOIL_Limitation_layers/finished_unmerged_layers/NZ-High-Alt-Treeline-layer-02_23_v2.gpkg
L3 : TOO_WET              : /DATA/SOIL_Limitation_layers/finished_unmerged_layers/LIMITING_SOILS_TOO_WET.shp
L4 : TOO_DRY              : /DATA/SOIL_Limitation_layers/finished_unmerged_layers/LIMITING_SOILS_TOO_DRY.shp
L5 : SALINE               : /DATA/SOIL_Limitation_layers/finished_unmerged_layers/SMAP_FSL_CMB_NZSC_FILT_SALINE.shp

Approach for merging a new layer Lb into the exisiting layer La

1. LaxLb: Get Intersection LaxLb. These are the polygons that will have a minimum of two limiting factors (at least one from La and exactly one from Lb)
2. La_minus_LaxLb: Get the difference between La and LaxLb. These are the polygons from the old layer that remain
3. Lb_minus_LaxLb: Get the difference between Lb and LaxLb. These are the polygons from the new layer that will be merged and extend the area of limitation


# MERGE_L0_L1 : PODZOLS (L1) ---> ELF_TO_KEEP_PROPER (L0)

```{r}

   LIMITING_SOILS_PODZOLS_ffn = paste0(elfdir,"DATA/SOIL_Limitation_layers/finished_unmerged_layers/LIMITING_SOILS_PODZOLS.shp")
LIMITING_SOILS_PODZOLS = st_read(LIMITING_SOILS_PODZOLS_ffn)

ELF_2012_TO_KEEP_PROPER_ffn = paste0(elf_merge_dir,"ELF_2012_TO_KEEP_PROPER.shp")
ELF_2012_TO_KEEP_PROPER = st_read(ELF_2012_TO_KEEP_PROPER_ffn)


  ffn_stub_list = c("L0xL1", "L0_minus_L0xL1", "L1_minus_L0xL1", "L0_L1_MERGED")
  LA_INPUT = ELF_2012_TO_KEEP_PROPER
  LB_INPUT = LIMITING_SOILS_PODZOLS
  LB_LIM_FIELDNAME = "PODZOLS"
  LB_LIMITATION_descr = "Substrate Podzols"

  MERGE_L0_L1_RES = merge_elf_layers(LA_INPUT, LB_INPUT, LB_LIM_FIELDNAME, LB_LIMITATION_descr, elf_merge_dir,  ffn_stub_list, ncore_rec = 36) 
  
```
 

# MERGE_L1_L2 : TREELINE (L2) ---> L0_L1_MERGED 

```{r}

   #Merge TREELINE into L0_L1_MERGED to create the merged layer "L1_L2_MERGED"

  ffn_stub_list = c("L1xL2", "L1_minus_L1xL2", "L2_minus_L1xL2", "L1_L2_MERGED")
  LA_INPUT = st_read(paste0(elf_merge_dir, "L0_L1_MERGED.gpkg"))
  LB_INPUT = st_read(paste0(elf_merge_dir, "NZ-High-Alt-Treeline-layer-02_23_v2.gpkg"))
  # mapview(LB_INPUT)
  LB_LIM_FIELDNAME = "TREELINE"
  LB_LIMITATION_descr = "Too Cold Near Treeline"

  MERGE_L1_L2_RES = merge_elf_layers(LA_INPUT, LB_INPUT, LB_LIM_FIELDNAME, LB_LIMITATION_descr, elf_merge_dir,  ffn_stub_list, ncore_rec = 64)

```
 
 

# MERGE_L2_L3 : WET (L3) ---> L1_L2_MERGED 

```{r}
#Merge TOO_WET into L1_L2_MERGED to create the merged layer "L2_L3_MERGED"
  
  ffn_stub_list = c("L2xL3", "L2_minus_L2xL3", "L3_minus_L2xL3", "L2_L3_MERGED")
  LA_INPUT = st_read(paste0(elf_merge_dir, "L1_L2_MERGED.gpkg"))
  LB_INPUT = st_read(paste0(elfdir,"DATA/SOIL_Limitation_layers/finished_unmerged_layers/SOIL_LIMITATIONS_TOO_WET.shp"))
  mapview(LB_INPUT)
  LB_LIM_FIELDNAME = "WET"
  LB_LIMITATION_descr = "Too Wet"

  MERGE_L2_L3_RES = merge_elf_layers(LA_INPUT, LB_INPUT, LB_LIM_FIELDNAME, LB_LIMITATION_descr, elf_merge_dir,  ffn_stub_list, ncore_rec = 36)

```
 

# MERGE_L3_L4 : DRY (L4) ---> L2_L3_MERGED 

```{r}
#Merge TOO_DRY into L2_L3_MERGED to create the merged layer "L3_L4_MERGED"
  
  ffn_stub_list = c("L3xL4", "L3_minus_L3xL4", "L4_minus_L3xL4", "L3_L4_MERGED")
  LA_INPUT = st_read(paste0(elf_merge_dir, "L2_L3_MERGED.gpkg"))
  
  # mapview(LB_INPUT)
  LB_LIM_FIELDNAME = "DRY"
  LB_LIMITATION_descr = "Too Dry"

  MERGE_L3_L4_RES = merge_elf_layers(LA_INPUT, LB_INPUT, LB_LIM_FIELDNAME, LB_LIMITATION_descr, elf_merge_dir,  ffn_stub_list, ncore_rec = 36)
  
  L3_L4_MERGED = st_read(paste0(elf_merge_dir, "L3_L4_MERGED.gpkg"))
  mapview(L3_L4_MERGED, zcol = "LIMITATION")
  

```
 

# MERGE_L4_L5 : SALINE (L5) ---> L3_L4_MERGED 

```{r}
#Merge SALINE into L3_L4_MERGED to create the merged layer "L4_L5_MERGED"
  
  ffn_stub_list = c("L4xL5", "L4_minus_L4xL5", "L5_minus_L4xL5", "L4_L5_MERGED")
  LA_INPUT = st_read(paste0(elf_merge_dir, "L3_L4_MERGED.gpkg"))
  
  # mapview(LB_INPUT)
  LB_LIM_FIELDNAME = "SALINE"
  LB_LIMITATION_descr = "Substrate Saline"

  MERGE_L4_L5_RES = merge_elf_layers(LA_INPUT, LB_INPUT, LB_LIM_FIELDNAME, LB_LIMITATION_descr, elf_merge_dir,  ffn_stub_list, ncore_rec = 36)

```
 
# Area Checks 
 
```{r}

# read the 2012 Layer
ELF_2012_ffn = paste0(elfdir, "data/mfe-environmental-limiting-factors-2012-SHP/environmental-limiting-factors-2012.shp")
ELF_2012 = st_read(ELF_2012_ffn) %>% 
  filter(!LIMITATION=="None")

ELF_2012_ng = ELF_2012 %>% ea_id_area(rem_geo = T) 
ELF_2012_ng_SMY = ELF_2012_ng %>% group_by(LIMITATION) %>% summarise(AREA_KHA = sum(POLY_AREA_HA)/1e3)

ELF_2012_ng %>% select(DRY, FROST_FLAT:SALT_SPRAY, WET:TREELINE, POLY_AREA_HA) %>% 
  gather(LIM, BX, -POLY_AREA_HA) %>% 
  filter(BX==1) %>% 
  group_by(LIM) %>% 
  summarise(AREA_KHA = sum(POLY_AREA_HA)/1e3)
  
  
#The new layers

LIMITING_SOILS_PODZOLS_ffn = paste0(elfdir,"DATA/SOIL_Limitation_layers/finished_unmerged_layers/LIMITING_SOILS_PODZOLS.shp")
LIMITING_SOILS_PODZOLS = st_read(LIMITING_SOILS_PODZOLS_ffn)
LCDB_TL = st_read(paste0(elf_merge_dir, "NZ-High-Alt-Treeline-layer-02_23_v2.gpkg"))
WET = st_read(paste0(elfdir, "DATA/SOIL_Limitation_layers/finished_unmerged_layers/SOIL_LIMITATIONS_TOO_WET.shp"))
SALINE = st_read(paste0(elfdir,"DATA/SOIL_Limitation_layers/finished_unmerged_layers/SMAP_FSL_CMB_NZSC_FILT_SALINE.shp"))
DRY = st_read(paste0(elfdir,"DATA/SOIL_Limitation_layers/finished_unmerged_layers/SOIL_LIMITATIONS_TOO_DRY.shp"))

#The Previous ELF Layer with the Limitation Layers to be replaced removed
ELF_2012_TO_KEEP_PROPER_ffn = paste0(elf_merge_dir,"ELF_2012_TO_KEEP_PROPER.shp")
ELF_2012_TO_KEEP_PROPER = st_read(ELF_2012_TO_KEEP_PROPER_ffn)


calc_tot_area(LIMITING_SOILS_PODZOLS)
calc_tot_area(LCDB_TL)
calc_tot_area(WET)
calc_tot_area(DRY)
calc_tot_area(SALINE)




# Check the first merged layer
# We should find that there are 1193 kha of poszols in this layer

L0_L1_MERGED = st_read(paste0(elf_merge_dir, "L0_L1_MERGED.gpkg"))
L1_L2_MERGED = st_read(paste0(elf_merge_dir, "L1_L2_MERGED.gpkg"))
L2_L3_MERGED = st_read(paste0(elf_merge_dir, "L2_L3_MERGED.gpkg"))
L3_L4_MERGED = st_read(paste0(elf_merge_dir, "L3_L4_MERGED.gpkg"))
L4_L5_MERGED = st_read(paste0(elf_merge_dir, "L4_L5_MERGED.gpkg"))

ELF__2023_PREFINAL = st_read(paste0("C:/Users/McMillanAn/OneDrive - MWLR/Projects/PRJ3597-MPI-ELF/data/ELF-LAYER-2023-PREFINAL/ELF_2023_PREFINAL_FULL_CLEAN_SMY_VALID.gpkg"))

show_areas = function(X){
  
  
LVLS = c("GUMLANDS", "PAKIHI", "PODZOLS", "ULTRAMAFIC", "COLD_BASIN", "FROST_FLAT", "TREELINE", "DRY", "SALT_SPRAY", "HOT_GEOTHM", "WET")

  
  # X = ELF_2012
  X_ng = X %>% 
    ea_id_area(rem_geo = T) %>% 
    group_by(LIMITATION) %>% 
    summarise(AREA_KHA = sum(POLY_AREA_HA)/1e3)
  
  X_ng2 = X %>% 
    ea_id_area(rem_geo = T) %>% 
    select(DRY, FROST_FLAT:SALT_SPRAY, WET:TREELINE, POLY_AREA_HA) %>% 
  gather(LIM, BX, -POLY_AREA_HA) %>% 
  filter(BX==1) %>% 
  group_by(LIM) %>% 
  summarise(AREA_KHA = sum(POLY_AREA_HA)/1e3)
  
  X_ng2$LIM = factor(X_ng2$LIM, levels = LVLS)
  
  print("AREAS GROUPED BY THE LIMITATION FIELD (including Multiple Limitations")
  print(X_ng)
  print("AREAS BY AREA OF EACH LIMITING VARIABLE (includes overlaps where multi-lims occur")
  print(X_ng2)
  
}

show_areas(ELF_2012)
show_areas(ELF_2012_TO_KEEP_PROPER)
show_areas(L0_L1_MERGED)
show_areas(L1_L2_MERGED)
show_areas(L2_L3_MERGED)
show_areas(L3_L4_MERGED)
show_areas(L4_L5_MERGED)

show_areas(ELF__2023_PREFINAL)



```





 
 
  