---
title: "ELF-clean-and-QAQC"
output: html_notebook
editor_options: 
  chunk_output_type: console
---


Last updated 29/06/23 06:28

# libraries and folders


```{r}




library(sf)
library(tidyverse)
# library(mapview)
library(cowplot)
library(lubridate)
# library(xlsx)
library(readxl)
library(rmapshaper)
library(mapview)
# library(rmapshaper)
library(plotly)
library(tictoc)
library(parallel)
library(doParallel)
library(terra)
library(openxlsx)
library(scales)
library(purrr)
library(furrr)
library(terra)
library(nngeo)

Current_PC <- Sys.info()["nodename"]

if (Current_PC == ""){
  root = "T:/Palmerston North/Projects A-E/Environmental Limiting Factors/ELF_Project/"
  datadir = paste0(root, "DATA/")
  plotdir = paste0(root, "plot/")


} else if (Current_PC == "EANZ-DT01-Linux") {
  os <- "linux"
  rootDrive <- "/media/andrew/HDD data/"
  EDrive <- "/media/andrew/E-DRIVE-SSD/"
  GeoSpatRoot = "C:/Users/McMillanAn/OneDrive - MWLR/NZ-Geospatial-Data/"
  
} else if (Current_PC == "EANZ-DT01") {
  os <- "win10"
  rootDrive <- "D:/"
  EDrive <- "E:/"
  GeoSpatRoot = "C:/Users/McMillanAn/OneDrive - MWLR/NZ-Geospatial-Data/" #check this!
  
} else if (Current_PC == "styx") {
  os <- "linux"
  rootDrive <- "/home/mcmillana/"
  elfdir = "/mnt/data/mcmillana/PRJ3597-MPI-ELF/"
  GeoSpatRoot = "/mnt/data/mcmillana/NZ-Geospatial-Data/"
  source("/home/mcmillana/code/R/utils/mcm_sp_utils.R") 
  
} else if (Current_PC == "L-8P30JW3") {
  os <- "win10"
  rootDrive <- "C:/Users/McMillanAn/OneDrive - MWLR"
  elfdir = "T:/Palmerston North/Projects A-E/Environmental Limiting Factors/ELF_Project/"
  GeoSpatRoot = "C:/Users/McMillanAn/OneDrive - MWLR/NZ-Geospatial-Data/"
   source("C:/Users/McMillanAn/OneDrive - MWLR/code/R/utils/mcm_sp_utils.R")  

} else {
  rootDrive <- "D:/"
}

elf_merge_dir = paste0(elfdir, "DATA/MERGING_ELF_REV/")

```


# read old and new elf layers and box
```{r}

gc()
ELF_2023_PREFINAL_ffn = paste0(elfdir, "ELF-LAYER-2023-PREFINAL/ELF_2023_PREFINAL.gpkg")
ELF_2023_PREFINAL = st_read(ELF_2023_PREFINAL_ffn)

ELF_2012_ffn = paste0(elfdir, "DATA/mfe-environmental-limiting-factors-2012-SHP/environmental-limiting-factors-2012.shp")
ELF_2012 = st_read(ELF_2012_ffn) %>% 
  filter(!LIMITATION=="None")


ELF_2012 %>% ea_id_area() %>% st_set_geometry(NULL) %>% group_by(LIMITATION) %>% summarise(
  MIN_AREA = min(POLY_AREA_HA),
  MEAN_POLY_AREA = mean(POLY_AREA),
  max_POLY_AREA = max(POLY_AREA))

ELF_2012 %>% ea_id_area() %>% st_set_geometry(NULL) %>% group_by(LIMITATION) %>% 
  summarise(length(which(POLY_AREA_HA < .1)))

# ELF_TEST_BOX_ffn = paste0(elf_merge_dir,"ELF_TEST_BOX.shp" )
ELF_TEST_BOX_ffn = paste0(elfdir,"ELF-LAYER-2023-PREFINAL/BIG_BOX.shp")
ELF_TEST_BOX = st_read(ELF_TEST_BOX_ffn)

ELF_2012_CLIP = st_crop(ELF_2012, ELF_TEST_BOX)
```

#Clean Clip

```{r}

ELF_2023_PREFINAL_CLIP = st_crop(ELF_2023_PREFINAL, ELF_TEST_BOX) %>% 
  mutate(LIM_TYPE=
           strtoi(
    paste0(
      as.character(DRY),as.character(FROST_FLAT),as.character(SALINE),as.character(ULTRAMAFIC),
      as.character(HOT_GEOTHM),as.character(PAKIHI),as.character(GUMLANDS),as.character(SALT_SPRAY),
      as.character(WET),as.character(COLD_BASIN),as.character(PODZOLS),as.character(TREELINE)
    ), base=2)
  ) %>% 
ea_id_area() %>% 
  filter(POLY_AREA_HA > 1)

mapview(ELF_2023_PREFINAL_CLIP, zcol = "LIM_TYPE")

ELF_2023_PREFINAL_CLIP_SMY = ELF_2023_PREFINAL_CLIP %>% 
   group_by(LIM_TYPE) %>% 
  summarise(across(c(1:14,16), first)) %>% 
  st_remove_holes(1.5*1e4) %>% 
  st_union(by_feature = T, is_coverage = T)
  
mapview(ELF_2023_PREFINAL_CLIP_SMY, zcol = "LIM_TYPE")

```


#=================================
# Final Cleaning and Validation
#=================================

# Step 1. Read the merged layer

```{r}
# ELF_2023_PREFINAL_ffn = paste0(elfdir, "ELF-LAYER-2023-PREFINAL/ELF_2023_PREFINAL.gpkg")


ELF_2023_PREFINAL_ffn = paste0(elfdir, "DATA/MERGING_ELF_REV/L4_L5_MERGED.gpkg") 
ELF_2023_PREFINAL = st_read(ELF_2023_PREFINAL_ffn)

ELF_2023_PREFINAL_shp_ffn = paste0(elfdir, "DATA/MERGING_ELF_REV/L4_L5_MERGED.shp") 
st_write(ELF_2023_PREFINAL, ELF_2023_PREFINAL_shp_ffn, append = F)


ELF_2023_PREFINAL_cln = ELF_2023_PREFINAL %>% 
  st_cast("POLYGON") %>% 
  st_collection_extract("POLYGON") %>% 
  ea_id_area() %>% 
  filter(POLY_AREA_HA < 1)


val = length(!st_is_valid(ELF_2023_PREFINAL_cln))


ELF_2023_PREFINAL_cln_val_01 = st_make_valid()




```

# Step 2 and 3. Add the LIM_TYPE Integer and filter out polygons < 1ha

```{r}
ELF_2023_PREFINAL_FULL_CLEAN = ELF_2023_PREFINAL %>% 
  mutate(LIM_TYPE=
           strtoi(
    paste0(
      as.character(DRY),as.character(FROST_FLAT),as.character(SALINE),as.character(ULTRAMAFIC),
      as.character(HOT_GEOTHM),as.character(PAKIHI),as.character(GUMLANDS),as.character(SALT_SPRAY),
      as.character(WET),as.character(COLD_BASIN),as.character(PODZOLS),as.character(TREELINE)
    ), base=2)
  ) %>% 
ea_id_area() %>% 
  filter(POLY_AREA_HA > 1)
```

# Step 4. save this new layer

```{r}
ELF_2023_PREFINAL_FULL_CLEAN_ffn = paste0(elfdir, "ELF-LAYER-2023-PREFINAL/ELF_2023_PREFINAL_FULL_CLEAN.gpkg") 
st_write(ELF_2023_PREFINAL_FULL_CLEAN, ELF_2023_PREFINAL_FULL_CLEAN_ffn, append = F)
```


# Step 5. Group the layer by the LIM_TYPE integer and summarise

```{r}
ELF_2023_PREFINAL_FULL_CLEAN_SMY = ELF_2023_PREFINAL_FULL_CLEAN %>% 
  group_by(LIM_TYPE) %>% 
  summarise(across(c(1:14,16), first))
```

# Step 6. Save the new summarised layer (ELF_2023_PREFINAL_FULL_CLEAN_SMY)

```{r}
ELF_2023_PREFINAL_FULL_CLEAN_SMY_ffn = paste0(elfdir, "ELF-LAYER-2023-PREFINAL/ELF_2023_PREFINAL_FULL_CLEAN_SMY.gpkg") 
st_write(ELF_2023_PREFINAL_FULL_CLEAN_SMY, ELF_2023_PREFINAL_FULL_CLEAN_SMY_ffn, append = F)
```

# Step 7. Fill any hole up to 1.5 ha then save as ELF-LAYER-2023-PREFINAL/ELF_2023_PREFINAL_FULL_CLEAN_SMY_HOLES_FILLED.gpkg

```{r}
ELF_2023_PREFINAL_FULL_CLEAN_SMY_HOLES_FILLED = ELF_2023_PREFINAL_FULL_CLEAN_SMY %>% 
  st_remove_holes(1.5*1e4) 

ELF_2023_PREFINAL_FULL_CLEAN_SMY_HOLES_FILLED_ffn = paste0(elfdir, "ELF-LAYER-2023-PREFINAL/ELF_2023_PREFINAL_FULL_CLEAN_SMY_HOLES_FILLED.gpkg") 
st_write(ELF_2023_PREFINAL_FULL_CLEAN_SMY_HOLES_FILLED, ELF_2023_PREFINAL_FULL_CLEAN_SMY_HOLES_FILLED_ffn, append = F)
```

read the layer from disk if reqd

```{r}
ELF_2023_PREFINAL_FULL_CLEAN_SMY_HOLES_FILLED_ffn = paste0(elfdir, "ELF-LAYER-2023-PREFINAL/ELF_2023_PREFINAL_FULL_CLEAN_SMY_HOLES_FILLED.gpkg") 

ELF_2023_PREFINAL_FULL_CLEAN_SMY_HOLES_FILLED = st_read(ELF_2023_PREFINAL_FULL_CLEAN_SMY_HOLES_FILLED_ffn)
```

# Step 8. Now validate the layer, cast as multipolygons and save as "ELF-LAYER-2023-PREFINAL/ELF_2023_PREFINAL_FULL_CLEAN_SMY_HOLES_FILLED_VALID.gpkg"

```{r}
n_valid_before = length(which(st_is_valid(ELF_2023_PREFINAL_FULL_CLEAN_SMY_HOLES_FILLED )))

ELF_2023_PREFINAL_FULL_CLEAN_SMY_HOLES_FILLED_VALID = ELF_2023_PREFINAL_FULL_CLEAN_SMY_HOLES_FILLED %>% 
  st_make_valid() %>% 
  st_cast("MULTIPOLYGON")

n_valid_after = length(which(st_is_valid(ELF_2023_PREFINAL_FULL_CLEAN_SMY_HOLES_FILLED_VALID )))


ELF_2023_PREFINAL_FULL_CLEAN_SMY_HOLES_FILLED_VALID_ffn = paste0(elfdir, "ELF-LAYER-2023-PREFINAL/ELF_2023_PREFINAL_FULL_CLEAN_SMY_HOLES_FILLED_VALID.gpkg") 
st_write(ELF_2023_PREFINAL_FULL_CLEAN_SMY_HOLES_FILLED_VALID, ELF_2023_PREFINAL_FULL_CLEAN_SMY_HOLES_FILLED_VALID_ffn, append = F)
```

# Step 9. Rename this layer and save to the final directory

```{r}

ELF_2023_PREFINAL_FULL_CLEAN_SMY_HOLES_FILLED_VALID_ffn = paste0(elfdir, "ELF-LAYER-2023-PREFINAL/ELF_2023_PREFINAL_FULL_CLEAN_SMY_HOLES_FILLED_VALID.gpkg") 

if (!exists("ELF_2023_PREFINAL_FULL_CLEAN_SMY_HOLES_FILLED_VALID")){ELF_2023_PREFINAL_FULL_CLEAN_SMY_HOLES_FILLED_VALID = st_read(ELF_2023_PREFINAL_FULL_CLEAN_SMY_HOLES_FILLED_VALID_ffn)}

ELF_2023_FINAL = ELF_2023_PREFINAL_FULL_CLEAN_SMY_HOLES_FILLED_VALID

ELF_2023_FINAL_dn = paste0(elfdir, "ELF-LAYER-2023-FINAL/")

ELF_2023_FINAL_SHP_dn = paste0(ELF_2023_FINAL_dn, "SHP/")
ELF_2023_FINAL_SHP_fn = "ELF_2023_FINAL.shp"
ELF_2023_FINAL_SHP_ffn = paste0(ELF_2023_FINAL_SHP_dn, ELF_2023_FINAL_SHP_fn)
if (!dir.exists(ELF_2023_FINAL_SHP_dn)){dir.create(ELF_2023_FINAL_SHP_dn, recursive = T)}
st_write(ELF_2023_FINAL, ELF_2023_FINAL_SHP_ffn, append = F)

ELF_2023_FINAL_GPKG_dn = paste0(ELF_2023_FINAL_dn, "GPKG/")
ELF_2023_FINAL_GPKG_fn = "ELF_2023_FINAL.gpkg"
ELF_2023_FINAL_GPKG_ffn = paste0(ELF_2023_FINAL_GPKG_dn, ELF_2023_FINAL_GPKG_fn)
if (!dir.exists(ELF_2023_FINAL_GPKG_dn)){dir.create(ELF_2023_FINAL_GPKG_dn, recursive = T)}
st_write(ELF_2023_FINAL, ELF_2023_FINAL_GPKG_ffn, append = F)

# The export to geodatabase can be done in QGIS

# ELF_2023_FINAL_GDB_dn = paste0(ELF_2023_FINAL_dn, "GDB/")
# ELF_2023_FINAL_GDB_fn = "ELF_2023_FINAL.gdb"
# ELF_2023_FINAL_GDB_ffn = paste0(ELF_2023_FINAL_GDB_dn, ELF_2023_FINAL_GDB_fn)
# if (!dir.exists(ELF_2023_FINAL_GDB_dn)){dir.create(ELF_2023_FINAL_GDB_dn, recursive = T)}
# st_write(ELF_2023_FINAL, ELF_2023_FINAL_GDB_ffn, append = F)



```

# Step 10. Cast the final layer as a single polygon

```{r}

ELF_2023_FINAL_dn = paste0(elfdir, "ELF-LAYER-2023-FINAL/")

ELF_2023_PREFINAL_FULL_CLEAN_SMY_HOLES_FILLED_VALID_ffn = paste0(elfdir, "ELF-LAYER-2023-PREFINAL/ELF_2023_PREFINAL_FULL_CLEAN_SMY_HOLES_FILLED_VALID.gpkg") 

if (!exists("ELF_2023_PREFINAL_FULL_CLEAN_SMY_HOLES_FILLED_VALID")){
  LF_2023_PREFINAL_FULL_CLEAN_SMY_HOLES_FILLED_VALID = st_read(ELF_2023_PREFINAL_FULL_CLEAN_SMY_HOLES_FILLED_VALID_ffn)
  }

ELF_2023_PREFINAL_FULL_CLEAN_SMY_HOLES_FILLED_VALID_SPOLY = ELF_2023_PREFINAL_FULL_CLEAN_SMY_HOLES_FILLED_VALID %>% 
  st_cast("POLYGON")

ELF_2023_FINAL_SPOLY = ELF_2023_PREFINAL_FULL_CLEAN_SMY_HOLES_FILLED_VALID_SPOLY

ELF_2023_FINAL_SPOLY_SHP_dn = paste0(ELF_2023_FINAL_dn, "SHP_SPOLY/")
ELF_2023_FINAL_SPOLY_SHP_fn = "ELF_2023_FINAL_SPOLY.shp"
ELF_2023_FINAL_SPOLY_SHP_ffn = paste0(ELF_2023_FINAL_SPOLY_SHP_dn, ELF_2023_FINAL_SPOLY_SHP_fn)
if (!dir.exists(ELF_2023_FINAL_SPOLY_SHP_dn)){dir.create(ELF_2023_FINAL_SPOLY_SHP_dn, recursive = T)}
st_write(ELF_2023_FINAL_SPOLY, ELF_2023_FINAL_SPOLY_SHP_ffn, append = F)


```

```{r}
mapview(ELF_2023_PREFINAL_FULL_CLEAN_SMY_HOLES_FILLED, zcol = "LIM_TYPE", alpha=0)

#----

mapview(ELF_2023_PREFINAL_FULL_CLEAN, zcol = "LIM_TYPE")


tic()
ELF_2023_PREFINAL_FULL_CLEAN_SMY = ELF_2023_PREFINAL_FULL_CLEAN %>% 
   group_by(LIM_TYPE) %>% 
  summarise(across(c(1:14,16), first)) %>% 
  st_remove_holes(1.5*1e4) 
toc()
# %>% 
#   st_union(by_feature = T, is_coverage = T)
  
ELF_2023_PREFINAL_FULL_CLEAN_SMY_ffn = paste0(elfdir, "ELF-LAYER-2023-PREFINAL/ELF_2023_PREFINAL_FULL_CLEAN_SMY.gpkg")

st_write(ELF_2023_PREFINAL_FULL_CLEAN_SMY, ELF_2023_PREFINAL_FULL_CLEAN_SMY_ffn, append = T)


ELF_2023_PREFINAL_FULL_CLEAN_SMY = st_read(ELF_2023_PREFINAL_FULL_CLEAN_SMY_ffn)

ELF_2023_PREFINAL_FULL_CLEAN_SMY_valid = st_make_valid(ELF_2023_PREFINAL_FULL_CLEAN_SMY) 

ELF_2023_PREFINAL_FULL_CLEAN_SMY_valid_POLY = ELF_2023_PREFINAL_FULL_CLEAN_SMY_valid %>% 
  st_cast("MULTIPOLYGON")
  

ELF_2023_PREFINAL_FULL_CLEAN_SMY_valid_POLY_ffn = paste0(elfdir, "ELF-LAYER-2023-PREFINAL/ELF_2023_PREFINAL_FULL_CLEAN_SMY_valid_POLY.gpkg")

st_write(ELF_2023_PREFINAL_FULL_CLEAN_SMY_valid_POLY, ELF_2023_PREFINAL_FULL_CLEAN_SMY_valid_POLY_ffn, append = F)







# length(which(!valid_bix))
# 
# ELF_2023_PREFINAL_FULL_CLEAN_SMY_valid_ffn = paste0(elfdir,"ELF-LAYER-2023-PREFINAL/ELF_2023_PREFINAL_FULL_CLEAN_SMY_valid.gpkg") 
# 
# st_write(ELF_2023_PREFINAL_FULL_CLEAN_SMY_valid, ELF_2023_PREFINAL_FULL_CLEAN_SMY_valid_ffn, append = F)
# mapview(ELF_2023_PREFINAL_FULL_CLEAN_SMY, zcol = "LIM_TYPE")

```


```{r}
ELF_2012 %>% pull(LIMITATION) %>% unique()



ELF_2012_CLIP_CMB = st_combine(ELF_2012_CLIP)
ELF_2023_PREFINAL_CLIP_CMB = st_combine(ELF_2023_PREFINAL_CLIP)

ELF_2012_CLIP_UNN = st_union(ELF_2012_CLIP)
ELF_2023_PREFINAL_CLIP_UNN = st_union(ELF_2023_PREFINAL_CLIP)

A = mapview(ELF_2012_CLIP, col.region = "blue")
B = mapview(ELF_2023_PREFINAL_CLIP, col.region = "orange")

A + B

C = mapview(ELF_2012_CLIP_CMB, col.region = "blue")
D = mapview(ELF_2023_PREFINAL_CLIP_CMB, col.region = "orange")

C + D

G = mapview(ELF_2012_CLIP_UNN, col.region = "blue")
H = mapview(ELF_2023_PREFINAL_CLIP_UNN, col.region = "orange")

G + H

ELF_2012_SYMDIFF_ELF_2023 = st_sym_difference(ELF_2012_CLIP, ELF_2023_PREFINAL_CLIP)

qm(ELF_2012_SYMDIFF_ELF_2023)
ggplot(ELF_2012_SYMDIFF_ELF_2023) + geom_sf()
plot(ELF_2012_SYMDIFF_ELF_2023)
mapview(ELF_2012_SYMDIFF_ELF_2023)

```

Try the contains_properly

```{r}
library(nngeo)

ELF_2023_PREFINAL_CLIP_w_AREA = ELF_2023_PREFINAL_CLIP %>% 
  ea_id_area()

ELF_2023_PREFINAL_CLIP_NO_SMALL = ELF_2023_PREFINAL_CLIP_w_AREA %>% 
  filter(POLY_AREA_HA > .5)

ELF_2023_PREFINAL_CLIP_NO_SMALL_FILLED_HOLES = st_remove_holes(ELF_2023_PREFINAL_CLIP_NO_SMALL, 9000)

A = mapview(ELF_2023_PREFINAL_CLIP_w_AREA, col.region = "grey")
B=mapview(ELF_2023_PREFINAL_CLIP_NO_SMALL, col.region = "green")
C=mapview(ELF_2023_PREFINAL_CLIP_NO_SMALL_FILLED_HOLES, col.region="orange")

C + B + A
```

# Create Difference Layer

```{r}

elf_final_dn = paste0(elfdir,"ELF-LAYER-2023-FINAL/")

ELF_2012_ffn = paste0(elfdir, "DATA/mfe-environmental-limiting-factors-2012-SHP/environmental-limiting-factors-2012.shp")

ELF_2012 = st_read(ELF_2012_ffn) %>% 
  filter(!LIMITATION=="None")
# 
# ELF_2023_ffn = paste0(elf_final_dn, "ELF_2023_FINAL.gpkg") 
# ELF_2023 = st_read(ELF_2023_ffn)

ELF_2023_PREFINAL_FULL_CLEAN_SMY_valid_POLY_ffn = paste0(elfdir, "ELF-LAYER-2023-PREFINAL/ELF_2023_PREFINAL_FULL_CLEAN_SMY_valid_POLY.gpkg")

ELF_2023 = st_read(ELF_2023_PREFINAL_FULL_CLEAN_SMY_valid_POLY_ffn)

ELF_2012_UNN = st_union(ELF_2012)
ELF_2023_UNN = st_union(ELF_2023)

ncore_rec = 64

  ELF_2012_minus_ELF_2023 = ea_intersection(ELF_2012, ELF_2023_UNN,PC = Current_PC, ffn_stub = "ELF_2012_minus_ELF_2023", dn = elf_final_dn, parallel = T, ncore = ncore_rec, val = "both",filetype = "SHP",diff_opt = T )

ELF_2023_minus_ELF_2012 = ea_intersection(ELF_2023, ELF_2012_UNN,PC = Current_PC, ffn_stub = "ELF_2023_minus_ELF_2012", dn = elf_final_dn, parallel = T, ncore = ncore_rec, val = "both",filetype = "SHP",diff_opt = T )



```

