---
title: "ELF-MERGE"
output: html_notebook
editor_options: 
  chunk_output_type: console
---

```{r}




library(sf)
library(tidyverse)
# library(mapview)
library(cowplot)
library(lubridate)
# library(xlsx)
library(readxl)
library(rmapshaper)
library(mapview)
# library(rmapshaper)
library(plotly)
library(tictoc)
library(parallel)
library(doParallel)
library(terra)
library(openxlsx)
library(scales)
library(purrr)
library(furrr)
library(terra)

Current_PC <- Sys.info()["nodename"]

if (Current_PC == ""){
  root = "T:/Palmerston North/Projects A-E/Environmental Limiting Factors/ELF_Project/"
  datadir = paste0(root, "DATA/")
  plotdir = paste0(root, "plot/")


} else if (Current_PC == "EANZ-DT01-Linux") {
  os <- "linux"
  rootDrive <- "/media/andrew/HDD data/"
  EDrive <- "/media/andrew/E-DRIVE-SSD/"
  GeoSpatRoot = "C:/Users/McMillanAn/OneDrive - MWLR/NZ-Geospatial-Data/"
  
} else if (Current_PC == "EANZ-DT01") {
  os <- "win10"
  rootDrive <- "D:/"
  EDrive <- "E:/"
  GeoSpatRoot = "C:/Users/McMillanAn/OneDrive - MWLR/NZ-Geospatial-Data/" #check this!
  
} else if (Current_PC == "styx") {
  os <- "linux"
  rootDrive <- "/home/mcmillana/"
  elfdir = "/mnt/data/mcmillana/PRJ3597-MPI-ELF/"
  GeoSpatRoot = "/mnt/data/mcmillana/NZ-Geospatial-Data/"
  source("/home/mcmillana/code/R/utils/mcm_sp_utils.R")  
  
} else if (Current_PC == "L-8P30JW3") {
  os <- "win10"
  rootDrive <- "C:/Users/McMillanAn/OneDrive - MWLR"
  elfdir = "T:/Palmerston North/Projects A-E/Environmental Limiting Factors/ELF_Project/"
  GeoSpatRoot = "C:/Users/McMillanAn/OneDrive - MWLR/NZ-Geospatial-Data/"
  source("T:/Palmerston North/Projects A-E/Environmental Limiting Factors/ELF_Project/DATA/SOIL_Limitation_layers/Rcode/mcm_sp_utils.R")  
} else {
  rootDrive <- "D:/"
}



```

# function: find_overlapping 
```{r}

find_overlapping = function(L1, L2, datadir,  ffn_stub,ncore_rec = 18){
  
  tic()
  P1_ORIG =  L1 %>%  mutate(P1_UID = 1:nrow(.))
  P2_ORIG =  L2 %>%  mutate(P2_UID = 1:nrow(.))
  
  
  
  P1 = P1_ORIG  %>% select(P1_UID, geometry)
  P2 = P2_ORIG %>% select(P2_UID, geometry)
  
  P1_x_P2 = st_intersects(P1,P2)
  length(P1_x_P2)
  
  P1_overlaps_ix = which(lengths(P1_x_P2) > 0)
  length(P1_overlaps_ix)
  P1_overlaps = P1[P1_overlaps_ix,]
  
  # tic()
  # P2_x_P1 = st_intersects(P2,P1_overlaps)
  # toc()
  # length(P2_x_P1)
  
  P2xP1_overlaps = ea_intersection(P2, P1_overlaps,PC = Current_PC, ffn_stub = ffn_stub, dn = datadir, parallel = T, ncore = ncore_rec, val = "both",filetype = "SHP",diff_opt = F )
  
  print(toc())
  
  return(P2xP1_overlaps)
}

```

LAYERS

L0 : ELF_2012_KEEP        : /DATA/SOIL_Limitation_layers/finished_unmerged_layers/ELF_2012_TO_KEEP_PROPER.shp
L1 : PODZOLS              : /DATA/SOIL_Limitation_layers/finished_unmerged_layers/LIMITING_SOILS_PODZOLS.shp
L2 : TREELINE             : /DATA/SOIL_Limitation_layers/finished_unmerged_layers/NZ-High-Alt-Treeline-layer-02_23_v2.gpkg
L3 : TOO_WET              : /DATA/SOIL_Limitation_layers/finished_unmerged_layers/LIMITING_SOILS_TOO_WET.shp
L4 : TOO_DRY              : /DATA/SOIL_Limitation_layers/finished_unmerged_layers/LIMITING_SOILS_TOO_DRY.shp
L5 : SALINE               : /DATA/SOIL_Limitation_layers/finished_unmerged_layers/SMAP_FSL_CMB_NZSC_FILT_SALINE.shp

Approach for merging a new layer Lb into the exisiting layer La

1. LaxLb: Get Intersection LaxLb. These are the polygons that will have a minimum of two limiting factors (at least one from La and exactly one from Lb)
2. La_minus_LaxLb: Get the difference between La and LaxLb. These are the polygons from the old layer that remain
3. Lb_minus_LaxLb: Get the difference between Lb and LaxLb. These are the polygons from the new layer that will be merged and extend the area of limitation

```{r}


LIMITING_SOILS_PODZOLS_ffn = paste0(elfdir,"DATA/SOIL_Limitation_layers/finished_unmerged_layers/LIMITING_SOILS_PODZOLS.shp")
LIMITING_SOILS_PODZOLS = st_read(LIMITING_SOILS_PODZOLS_ffn)

ELF_2012_TO_KEEP_PROPER_ffn = paste0(elfdir, "DATA/MERGING_ELF/ELF_2012_TO_KEEP_PROPER.shp")
ELF_2012_TO_KEEP_PROPER = st_read(ELF_2012_TO_KEEP_PROPER_ffn)



L0 = ELF_2012_TO_KEEP_PROPER
L1 = LIMITING_SOILS_PODZOLS

L0xL1 = find_overlapping(L0,L1, datadir = paste0(elfdir, "DATA/MERGING_ELF/"), ffn_stub = "L0xL1",ncore_rec = 72) 

#515 sec on laptop 18 cores
#248 seconds on Styx 18 cores
#188 seconds on Styx 36 cores
#175 seconds on Styx 72 cores

LO_minus_L0xL1 = st_difference(L0, st_union(L0xL1))
L1_minus_L0xL1 = st_difference(L1, st_union(L0xL1))


```

