---
title: "R Notebook"
output: html_notebook
editor_options: 
  chunk_output_type: console
---




# Glossary

CLTL = Cieraard Latitude Treeline
LCDB = MWLR's Land Cover Database


# Libraries and folders


```{r}


library(tidyverse)
library(rjson)
library(sf)
library(mapview)
library(terra)
library(tictoc)
source("D:/HerronK/ELF_Project/AndrewsWork/MyCode/R/elf/ea_utils.R")

#set up the data folders
rdn = "D:/HerronK/ELF_Project/"
CLTL_rdn = paste0(rdn, "ELFData/Cieraad/")
AL_dn = paste0(rdn,"AndrewsWork/layers/")
NTL_dn = paste0(AL_dn,"New-Treeline-Layers/")


TL_classes_unfilt = c(14,15,16,43,55)


```



```{r}



```


# Step 1. get the CLTL

```{r Step 1}

#set up the data folders
rdn = "D:/HerronK/ELF_Project/"
CLTL_rdn = paste0(rdn, "ELFData/Cieraad/")
AL_dn = paste0(rdn,"AndrewsWork/layers/")

#read the Cieraad Treeline based on the original regression
CLTL_ffn = paste0(CLTL_rdn,"Vector-TL-Cieraad/out.shp")
CLTL = st_read(CLTL_ffn)

#read the Cieraad Treeline based on the original regression adjusted by 100m for the NI
CLTL_NI_100m_ffn = paste0(CLTL_rdn,"NI_100m_vector_Cieraad/out.shp")
CLTL_NI_100m = st_read(CLTL_NI_100m_ffn) %>% 
  mutate(Island = "NI")

#read a polygon for the North Island
NI_poly_ffn = paste0(AL_dn,"NI.gpkg")
NI_poly = st_read(NI_poly_ffn)

#read a polygon for the North Island
SI_poly_ffn = paste0(AL_dn,"SI.gpkg")
SI_poly = st_read(SI_poly_ffn)

#grab the original tree line for the SI only
CLTL_SI =st_intersection(CLTL, SI_poly) %>% 
  mutate(Island = "SI") %>% 
  select(DN, Island)

# mapview(SI_poly)
# mapview(CLTL)
# mapview(CLTL_SI)

#combine it with the adjusted treeline for the North Is
CLTL_adj = CLTL_SI %>% bind_rows(CLTL_NI_100m)
mapview(CLTL_adj)

#save the combined/NI-Adjusted tree line
CLTL_adj_ffn = paste0(AL_dn,"CLTL_adj.gpkg")
st_write(CLTL_adj, CLTL_adj_ffn, append = F)
st_read(CLTL_adj_ffn)

```



#  Step 2. Now intersect the CLTL_adj with LCDB

```{r}

LCDB5_ffn = "D:/HerronK/ELF_Project/ELFData/LCDB/lcdb-v50-land-cover-database-version-50-mainland-new-zealand.gpkg"
LCDB5 = st_read(LCDB5_ffn)

CLTL_adj_ffn = paste0(AL_dn,"CLTL_adj.gpkg")
CLTL_adj = st_read(CLTL_adj_ffn)

tic("Intersecting LCDB5 and CLTL_adj")
LCDBxCLTL_adj = ea_intersection(LCDB5, CLTL_adj, PC = "PNGeo04-W10", "LCDBxCLTL_adj", dn = AL_dn, val = "none", parallel = T, ncore = 7, filetype = "GPKG", diff_opt = F, simplify_opt = F, tol = NULL)
toc()


LCDBxCLTL_adj_ffn = paste0(AL_dn, "LCDBxCLTL_adj.gpkg")
st_write(LCDBxCLTL_adj, LCDBxCLTL_adj_ffn, append = F)  

```

#  Step 3. Get diff between clipped features and LCDB features that intersect with CLTL

```{r}


#Define the full set of treeline classes



if (!exists("LCDBxCLTL_adj")){
  LCDBxCLTL_adj_ffn = paste0(AL_dn, "LCDBxCLTL_adj.gpkg")
  LCDBxCLTL_adj = st_read(LCDBxCLTL_adj_ffn)  
}

if (!exists("LCDB5")){
  LCDB5_ffn = "D:/HerronK/ELF_Project/ELFData/LCDB/lcdb-v50-land-cover-database-version-50-mainland-new-zealand.gpkg"
  LCDB5 = st_read(LCDB5_ffn)
  
}

# get the LCDB features that intersect with the CTLT
intersecting_features = LCDBxCLTL_adj %>% pull(LCDB_UID) %>% unique()
LCDB5_intersecting_feat = LCDB5 %>% 
  filter(LCDB_UID %in% intersecting_features & Class_2018 %in% TL_classes_unfilt)

mapview(LCDB5_intersecting_feat)

# now create a layer of the split features that lie outside of the CLTL (takes 7 minutes sec)
tic("Getting difference between LCDB5_intersecting_feat and LCDBxCLTL_adj")
LCDB5_intersecting_feat_minus_LCDBxCLTL_adj = getLayerDiff(LCDB5_intersecting_feat, LCDBxCLTL_adj)
toc()

#add in the island information

island_key = LCDBxCLTL_adj %>% 
  st_set_geometry(NULL) %>% 
  select(LCDB_UID, Island) %>% 
  distinct() %>% 
  as_tibble()

LCDB5_intersecting_feat_minus_LCDBxCLTL_adj = LCDB5_intersecting_feat_minus_LCDBxCLTL_adj %>% 
  left_join(island_key, by = "LCDB_UID")


LCDB5_intersecting_feat_minus_LCDBxCLTL_adj_ffn = paste0(AL_dn, "New-Treeline-Layers/LCDB5_intersecting_feat_minus_LCDBxCLTL_adj.gpkg")
st_write(LCDB5_intersecting_feat_minus_LCDBxCLTL_adj, LCDB5_intersecting_feat_minus_LCDBxCLTL_adj_ffn,append = F)



```

#  Step 4. extract DEM zonal stats from outside treeline class polys

```{r}

library(exactextractr)
# 
# # load the layer
# if (!exists("LCDBxCLTL_adj")){
#   LCDBxCLTL_adj_ffn = paste0(AL_dn, "LCDBxCLTL_adj.gpkg")
#   LCDBxCLTL_adj = st_read(LCDBxCLTL_adj_ffn)  
# }

LCDB5_intersecting_feat_minus_LCDBxCLTL_adj_ffn = paste0(AL_dn, "New-Treeline-Layers/LCDB5_intersecting_feat_minus_LCDBxCLTL_adj.gpkg")
LCDB5_intersecting_feat_minus_LCDBxCLTL_adj = st_read(LCDB5_intersecting_feat_minus_LCDBxCLTL_adj_ffn)

outside_polys = LCDB5_intersecting_feat_minus_LCDBxCLTL_adj

# grab the dem

dem_ffn = "D:/HerronK/ELF_Project/ELFData/NZ_DEM/dem_linz_nz_110520_orthonzgd49h_nztm_10m.kea"
dem = rast(dem_ffn)

# extract zonal stats from the DEM

outside_polys$alt_stats = exactextractr::exact_extract(dem, outside_polys, c("min", "mean"))

outside_polys_w_ZSTATS = outside_polys %>% 
  mutate(min_alt = alt_stats$min, mean_alt = alt_stats$mean) %>% 
  select(-contains("stats"))



names(LCDBxCLTL_adj)  
names(outside_polys_w_ZSTATS)  
# save the data with the Zonal Stats

# LCDBxCLTL_adj_w_ZSTATS_ffn = paste0(AL_dn, "LCDBxCLTL_adj_w_ZSTATS_v02.gpkg")
# st_write(LCDBxCLTL_adj_w_ZSTATS, LCDBxCLTL_adj_w_ZSTATS_ffn, append = F)


outside_polys_w_ZSTATS_ffn = paste0(AL_dn, "outside_polys_w_ZSTATS.gpkg")
st_write(outside_polys_w_ZSTATS, outside_polys_w_ZSTATS_ffn, append = F)

names(outside_polys_w_ZSTATS)

```


#  Step 5. Filter the outside_polys

apply the rules set out in Section 3.2ii of Caitlin's Report

```{r}

outside_polys_w_ZSTATS_ffn = paste0(AL_dn, "outside_polys_w_ZSTATS.gpkg")
outside_polys_w_ZSTATS = st_read(outside_polys_w_ZSTATS_ffn)


#Class 16 (Gravel and Rock) Rules
# 1. Any Class 16 Features that lie fully or mostly within layer are kept
# 2. Any Class 16 Features that intersect with layer but are mostly outside CLTL
# 1. Any Class 16 Features that lie within layer are kept

# #check that LCDB_UID is indeed a unique ID --> CONFIRMED
# (LCDB5 %>% pull(LCDB_UID) %>% unique() %>% length()) == LCDB5 %>% nrow()
# 
# #get the original areas of each polygon from LCDB
# LCDB5_orig_areas = LCDB5 %>% 
#   mutate(ORIG_AREA_ha = as.numeric(st_area(.))/1e4) %>% 
#   st_set_geometry(NULL) %>% 
#   select(LCDB_UID, ORIG_AREA_ha)


#for all the intersecting polygons add the orginal area
outside_polys_w_ZSTATS_mdf = 
  outside_polys_w_ZSTATS %>% mutate(AREA_ha = as.numeric(st_area(.))/1e4) 



#create flags:
# TREELINE_CLASS_FLAG: if class is 14,15,43,55
# CLASS 14: Permanent Snow and Ice
# CLASS 15: Alpine grass / herbfield
# CLASS_16: if class is 16 (gravel and rock)
# CLASS_43_OR_55_FLAG (43 = Tall Tussuck, 55 = Sub-Alpine Scrubland)
# CLASS_43_OR_55_GOOD_ALT ()

FRAC_INSIDE_THRESH = 0.5

#Apply the Filtering Logic on P3-4 of ETL report

outside_polys_flagged = outside_polys_w_ZSTATS_mdf %>% 
  #flag the class 16 data that mostly lie outside CLTL
  mutate(
    TREELINE_CLASS_FLAG = ifelse(Class_2018 %in% TL_classes_unfilt, TRUE, FALSE),
    KEEP_FLAG = case_when(
      # Not a treeline class -> Reject
      !TREELINE_CLASS_FLAG ~ "REJECT",
      # Class 14 and 15 are automatically kept
      Class_2018 == 14 | Class_2018 == 15 ~ "KEEP",
      # Class 16 features mostly outside and original area is > 10 ha need to be checked
      (Class_2018 == 16) & (AREA_ha >= 10)  ~ "CHECK",
      (Class_2018 == 16) & (AREA_ha < 10)  ~ "KEEP",
      # (Class_2018 != 16) & (FRAC_INSIDE >= FRAC_INSIDE_THRESH) ~ "N/A",
      # Class 43 and 55
      (Class_2018 == 43 | Class_2018 == 55) & (Island == "NI") & (min_alt > 900 | mean_alt > 1200) ~ "KEEP",
      (Class_2018 == 43 | Class_2018 == 55) & (Island == "NI") & (min_alt <= 900 & mean_alt <= 1200) ~ "CHECK",
      (Class_2018 == 43 | Class_2018 == 55) & (Island == "SI") &  (mean_alt > 1300) ~ "KEEP",
      (Class_2018 == 43 | Class_2018 == 55) & (Island == "SI") &   mean_alt > 1200 & min_alt > 400 ~ "KEEP",
      (Class_2018 == 43 | Class_2018 == 55) & (Island == "SI") &  (mean_alt <= 1200 | min_alt <= 400) ~ "CHECK",
      (Class_2018 == 43 | Class_2018 == 55) & (Island == "SI") &  (mean_alt < 1300 & (!(mean_alt > 1200 & min_alt > 400))) ~ "CHECK"))

outside_polys_flagged_ffn = paste0(AL_dn,"New-Treeline-Layers/outside_polys_flagged.gpkg")
st_write(outside_polys_flagged, outside_polys_flagged_ffn, append = F)


#How many require checking?

outside_polys_flagged %>% nrow() #17408
outside_polys_flagged %>% filter(KEEP_FLAG == "KEEP") %>% nrow() #6722
outside_polys_flagged %>% filter(KEEP_FLAG == "CHECK") %>% nrow() #10686
outside_polys_flagged %>% filter(KEEP_FLAG == "CHECK" | KEEP_FLAG == "KEEP") %>% nrow() #17408
outside_polys_flagged %>% filter(KEEP_FLAG == "REJECT") %>% nrow()  # 0


#of the checked polys what are the classes

outside_polys_flagged %>% 
  st_set_geometry(NULL) %>% 
  as_tibble() %>% 
  filter(KEEP_FLAG == "CHECK") %>% 
  group_by(Class_2018) %>% 
  summarise(
    Num = length(Class_2018),
    Area_avg = mean(AREA_ha),
    mean_alt_avg = mean(mean_alt),
    mean_alt_sd = sd(mean_alt),
    mean_alt_ln = length(mean_alt),
    mean_alt_min = min(mean_alt),
    mean_alt_max = max(mean_alt),
    mean_alt_med = median(mean_alt))




```

#  Step 6. Grab the Clipped Layer and then combine it with the filtered outsidepolys


```{r}

outside_polys_flagged_ffn = paste0(AL_dn,"New-Treeline-Layers/outside_polys_flagged.gpkg")
outside_polys_flagged = st_read(outside_polys_flagged_ffn)

LCDBxCLTL_adj_ffn = paste0(AL_dn, "LCDBxCLTL_adj.gpkg")
LCDBxCLTL_adj = st_read(LCDBxCLTL_adj_ffn)  

#Temporary - for now - keep only those flagged with keep

outside_polys_filtered = outside_polys_flagged %>% 
  filter(KEEP_FLAG == "KEEP")


outside_polys_filtered_UNN_ffn = paste0(NTL_dn, "outside_polys_filtered_UNN.gpkg")
st_write(outside_polys_filtered_UNN, outside_polys_filtered_UNN_ffn, append = F)

outside_polys_filtered_UNN_ffn = paste0(NTL_dn, "outside_polys_filtered_UNN.gpkg")
st_write(outside_polys_filtered_UNN, outside_polys_filtered_UNN_ffn, append = F)


inside_polys_filtered_UNN = LCDBxCLTL_adj %>% 
  filter(Class_2018 %in% TL_classes_unfilt) %>% 
  validate_sh() %>% 
  group_by(LCDB_UID) %>% 
  st_union(by_feature = T)

inside_polys_filtered_UNN_ffn = paste0(NTL_dn, "inside_polys_filtered_UNN.gpkg")
st_write(inside_polys_filtered_UNN, inside_polys_filtered_UNN_ffn, append = F)



#-----------------------testing how to do spatial join by field -----#
# define a test area

TB = st_read(paste0(AL_dn, "testbox2.shp"))

outside_polys_filtered_UNN_clip = st_intersection(outside_polys_filtered_UNN, TB)
inside_polys_filtered_UNN_clip = st_intersection(inside_polys_filtered_UNN, TB)




all_polys_filtered_UNN_clip = outside_polys_filtered_UNN_clip %>% 
  bind_rows(inside_polys_filtered_UNN_clip) %>% 
  group_by(LCDB_UID) %>% 
  summarise(LCDB_UID_first = first(LCDB_UID))

C = mapview(all_polys_filtered_UNN_clip)

outside_polys_filtered_UNN_clip$mean_alt

A = mapview(outside_polys_filtered_UNN_clip)
B = mapview(inside_polys_filtered_UNN_clip, col.region = "pink")

B + A

#------------- end of test ---------#


# combine the polys in and out then spatially join the polys with the same UID
# this takes about 10 minutes. You also lose all the data except for the geometry
# we get 35619 features

tic()
all_polys_filtered = inside_polys_filtered_UNN %>% 
  bind_rows(outside_polys_filtered_UNN) %>% 
  group_by(LCDB_UID) %>% 
  summarise(LCDB_UID_first = first(LCDB_UID))
toc()

#write this new layer to disk
all_polys_filtered_ffn = paste0(NTL_dn, "all_polys_filtered.gpkg")
st_write(all_polys_filtered, all_polys_filtered_ffn)


LCDBxCLTL_adj_nogeo = LCDBxCLTL_adj %>% 
  st_set_geometry(NULL) %>% 
  as_tibble()

#join in the data
all_polys_filtered_mdf = all_polys_filtered %>% 
  left_join(LCDBxCLTL_adj_nogeo, by = "LCDB_UID")

# tic()
# all_polys_filtered_UNN = 
#   all_polys_filtered %>% 
#   group_by(LCDB_UID) %>% 
#   summarise(Class_2018_first = first(Class_2018)) %>% 
#   st_cast()
# toc()





mapview(outside_polys_filtered_CMB)

outside_polys_filtered_UNN = st_union(validate_sh(outside_polys_filtered))

outside_polys_filtered_UNN_ffn = paste0(NTL_dn, "outside_polys_filtered_UNN.gpkg")
st_write(outside_polys_filtered_UNN, outside_polys_filtered_UNN_ffn, append = F)

mapview(outside_polys_filtered_UNN)

LCDBxCLTL_adj_CMB = st_combine(LCDBxCLTL_adj)


ATL_POLYS = LCDBxCLTL_adj %>% 
  st_combine(outside_polys_filtered)
  

```

```{r}
# New_TL_unfilt_w_FLAGS_ffn2 = paste0(AL_dn,"New-Treeline-Layers/New_TL_unfilt_w_FLAGS.shp")
# st_write(New_TL_unfilt_w_FLAGS, New_TL_unfilt_w_FLAGS_ffn2)

#Save the layer to check

New_TL_filtered_to_CHECK = New_TL_unfilt_w_FLAGS %>% 
  filter(KEEP_FLAG == "CHECK")

New_TL_filtered_to_CHECK_ffn = paste0(AL_dn,"New-Treeline-Layers/New_TL_filtered_to_CHECK.gpkg")
st_write(New_TL_filtered_to_CHECK, New_TL_filtered_to_CHECK_ffn, append = F)


#Create a new layer with just the shapes to check

New_TL_filtered_to_CHECK





```



#Test Extract

```{r}


dn = "D:/HerronK/ELF_Project/Taranaki/TaranakiProfiles/"
fn = "taranaki2_chmProfile.json"
ffn = paste0(dn,fn)

dn = "D:/HerronK/ELF_Project/gw/gwProfiles/"
fn = "gw6_chmProfile.json"
ffn = paste0(dn,fn)


file.exists(ffn)

prof = fromJSON(file = ffn) # %>% map_df(as_tibble)
prof[[1]]

df <- data.frame(matrix(unlist(prof), ncol = max(lengths(prof))+1, byrow = TRUE)) %>% as_tibble()
nms = c("dist",   "coords_1", "coords_2", "Min",    "Mean",   "Median", "Max",    "Std"   )

names(df) = nms


g = ggplot(df) + geom_line(aes(dist, Mean))
g

df_mdf = df %>% 
  mutate(
    coords_1_diff = c(0,diff(coords_1)),
    coords_2_diff = c(0,diff(coords_2)),
    coords_1_diff_sq = coords_1_diff^2,
    coords_2_diff_sq = coords_2_diff^2,
    coords_dist = sqrt(coords_1_diff_sq + coords_2_diff_sq))




coords_tr = df_mdf %>% select(lat = coords_1, lon = coords_2)

coords_sf = st_as_sf(coords_tr, coords = c("lat", "lon"), crs = 2193)
mapview(coords_sf)

coords_shp_ffn =  "D:/HerronK/ELF_Project/AndrewsWork/TestTr.shp"
st_write(coords_sf , coords_shp_ffn)



```

# Intersect LCDB with the Masked DEM Layer 

```{r}

library(tidyverse)
library(sf)
library(mapview)
library(terra)

source("D:/HerronK/ELF_Project/AndrewsWork/MyCode/R/ea_utils.R")

LCDB5_ffn = "D:/HerronK/ELF_Project/ELFData/LCDB/lcdb-v50-land-cover-database-version-50-mainland-new-zealand.gpkg"
file.exists(LCDB5_ffn)

# dem_dn = "D:/HerronK/ELF_Project/ELFData/NZ_DEM/"
# dem_fn = "dem_linz_nz_110520_orthonzgd49h_nztm_10m.kea"
# dem_ffn = paste0(dem_dn, dem_fn)
# dem = rast(dem_ffn)
# plot(dem)
# class(dem)

root1 = "D:/HerronK/ELF_Project/AndrewsWork/"
DEM_masked = st_read(paste0(root1, "DEM_MASKED_POLY.json"))

mapview(DEM_masked)


# A = st_intersection("")

DIROUT = "D:/HerronK/ELF_Project/AndrewsWork/"

DEM_maskedxLCDB = ea_intersection(LCDB5, DEM_masked, PC = "PNGeo04-W10", "Stub_", dn = DIROUT, val = "none", parallel = T, ncore = 7, filetype = "GPKG", diff_opt = F, simplify_opt = F, tol = NULL)



treeline_classes_init = c(14,15,16,43,55)
treeline_init = LCDB5 %>% filter(Class_2018 %in% treeline_classes_init)

mapview(treeline_init)

CLT

```

# Make a shape file of all the available LIDAR tiles

```{r}

dn = "D:/HerronK/ELF_Project/AndrewsWork/Lidar-Index-Tiles-LINZ/"


lidar_tiles_dn = list.dirs(dn, recursive = F)

for (i in 1:length(lidar_tiles_dn)){
  
  # i = 2
  
  cdir = lidar_tiles_dn[i]
  fls = list.files(cdir, full.names = T)
  
  
  ix = which(str_detect(fls, ".shp"))
  cfile = fls[ix]
  
  SF = st_read(cfile)
  
  name_tileset = str_extract(basename(cfile),  ".+?(?=-lidar)")
  print("************************************************************************************")
  print(paste("No.", i, name_tileset))
  print("************************************************************************************")
  print(paste("No.", i, name_tileset))
  SF_w_name = SF %>% mutate(name = name_tileset)
  if (i==1){
    SF_COL = SF_w_name
  }else{
    
    SF_COL = SF_COL %>% bind_rows(SF_w_name)
  }
  
  
  
  
  
  
  
  
  
  
}


NZ_LINZ_LIDAR_INDEX_TILES_ffn = paste0(dn, "All_LINZ_LiDAR_Index_Tiles.shp")
st_write(SF_COL, NZ_LINZ_LIDAR_INDEX_TILES_ffn, append = F)



```

# Dissolve internal lines in LIDAR INDEX TILES just to get outline

```{r}

NZ_LINZ_LIDAR_INDEX_TILES_ffn = paste0(dn, "All_LINZ_LiDAR_Index_Tiles.shp")
LIT = st_read(NZ_LINZ_LIDAR_INDEX_TILES_ffn)

LIT_sel = LIT %>% filter(name=="otago")

mapview(LIT_sel)


```

```{r}

franconia %>% 
  group_by(district)

mapview(franconia)

franconia %>% 
  group_by(district) %>%
  st_union() %>%
  mapview(zcol = "district")

franconia %>% 
  group_by(district) %>% 
  summarise(m = mean(SHAPE_LEN)) %>% 
  st_cast() %>% 
  mapview(zcol = "m")

```

